---
title: "Week One In-class Exercise"
output: html_notebook
---

# Anscombe Data Set

The Anscombe data set is indroduced in the *NIST Engineering and Statistics Handbook* to demonstrate the value of Exporatory Data Analysis (EDA)

```{r}

library(readxl)
Anscombe <- read_excel("~/Google Drive/UU PMST/MST 6600 - Advanced Statistical Techniques/NIST Engineering Statistics Handbook 1-1-6 EDA Graphics Example/data/Anscombe.xlsx")
```
Although this looks like a lot of code, it was generated using the **Import Dataset** dropdown box.

We can view the data in the concole, "viewer, " or in the table below:

```{r}
Anscombe
```


***

## What's intersting about this data set?

Given we have a set of X and Y pairs, we may assume that data should be grouped together; a normal, *non-graphical* analysis would be to preform a linear regression on the data.

```{r}
xy1.linear.model <- lm(Anscombe$Y1 ~ Anscombe$X1)
summary(xy1.linear.model)
```

I might even plot the data (adding the linear regression line from above):
```{r}
plot(Anscombe$X1, Anscombe$Y1)
abline(xy1.linear.model)
```

### Create summaries of the other XY pairs

```{r}
xy2.linear.model <- lm(Anscombe$Y2 ~ Anscombe$X2)
xy3.linear.model <- lm(Anscombe$Y3 ~ Anscombe$X3)
xy4.linear.model <- lm(Anscombe$Y4 ~ Anscombe$X4)
```

```{r}
summary(xy2.linear.model)
summary(xy3.linear.model)
summary(xy4.linear.model)
```

### Let's look at all the data graphically

```{r}
# create a single page of all values
par(mfrow=c(2,2)) # Change the panel layout to 2 x 2
plot(Anscombe$X1, Anscombe$Y1)
abline(xy1.linear.model)

plot(Anscombe$X2, Anscombe$Y2)
abline(xy2.linear.model)

plot(Anscombe$X3, Anscombe$Y3)
abline(xy3.linear.model)

plot(Anscombe$X4, Anscombe$Y4)
abline(xy4.linear.model)

par(mfrow=c(1,1)) # Change back to 1 x 1
```

### Let's make one of the graphs "pretty"
```{r}
# Making changes to graphs
# create a simple scatter plot of X1 vs Y1 with some lables
plot(Anscombe$X1, Anscombe$Y1, main="DATA SET 1", xlab="X1", ylab="Y1")
abline(xy1.linear.model)
```

### What else do I get with R?
```{r}
plot(Anscombe$X1, resid(xy1.linear.model))
abline(h = 0)

plot(Anscombe$X2, resid(xy2.linear.model))
abline(h = 0)
```
***

# A Look Ahead
The code above uses *base R* commands. While they can be used to create clean, accurate graphs, it requires the user to **know** many different commands and structures.

Over the last 10-years, several new R packages have been created to make analysis and graphics easier. This course will primarily exploit two of those packages: *tidyverse* and *ggplot2*.

## Example 2. The Anscombe dataset with ggplot2
I need to tidy up the data:

```{r}
library(tidyverse)
```

```{r}
Anscombe.tidy.xy1 <- Anscombe %>% select("X1", "Y1") %>%
  mutate(dataset = "DataSet1") %>%
  rename(X = X1, Y = Y1)

Anscombe.tidy.xy2 <- Anscombe %>% select("X2", "Y2") %>%
  mutate(dataset = "DataSet2") %>%
  rename(X = X2, Y = Y2)

Anscombe.tidy.xy3 <- Anscombe %>% select("X3", "Y3") %>%
  mutate(dataset = "DataSet3") %>%
  rename(X = X3, Y = Y3)

Anscombe.tidy.xy4 <- Anscombe %>% select("X4", "Y4") %>%
  mutate(dataset = "DataSet4") %>%
  rename(X = X4, Y = Y4)
```
We can look at the output of the first data set:
```{r}
Anscombe.tidy.xy1
```

In order to fully complete the analysis, we need to put all for *datasets* together:

```{r}
Anscombe.tidy.data <- 
  bind_rows(Anscombe.tidy.xy1, Anscombe.tidy.xy2, Anscombe.tidy.xy3, Anscombe.tidy.xy4)
```

The new dataset is "tidy."
```{r}
Anscombe.tidy.data
```

## The POWER of ggplot2

```{r}
Anscombe.tidy.data %>% 
  ggplot(aes(X, Y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ dataset) +
  theme_bw()
```


## Modeling is easier as well...
```{r}
# modeling the data for DataSet1 through DataSet4 
xy1.lm <- Anscombe.tidy.data %>% 
  filter(dataset == "DataSet1") %>% lm(Y ~ X, .)
#plot(xy1.lm)
summary(xy1.lm)

xy2.lm <- Anscombe.tidy.data %>% 
  filter(dataset == "DataSet2") %>% lm(Y ~ X, .)
#plot(xy2.lm)
summary(xy2.lm)

xy3.lm <- Anscombe.tidy.data %>% 
  filter(dataset == "DataSet3") %>% lm(Y ~ X, .)
#plot(xy3.lm)
summary(xy3.lm)

xy4.lm <- Anscombe.tidy.data %>% 
  filter(dataset == "DataSet4") %>% lm(Y ~ X, .)
#plot(xy4.lm)
summary(xy4.lm)

```